MG7: Configurable and scalable 16S metagenomics data analysis -- new methods optimized for massive cloud computing

### 2015-09-14 docs/paper/out/paper.tex
add Results Overview from @rtobes -- paper
see https://github.com/ohnosequences/metagenomica/issues/26#issuecomment-140027234

Considering all these aspects we have designed a new open source
methodology for analyzing metagenomics data that exploits the new
possibilities that cloud computing offers to get a system robust,
programmatically configurable, modular, distributed, flexible, scalable
and traceable in which the biological databases of reference sequences
can be easily updated and/or frequently substituted by new ones or by
databases specifically designed for focused projects.

@@

To tackle the challenges posed by metagenomics big data analysis
outlined in the Introduction

\begin{itemize}
\tightlist
\item
  Static reproducible specification of dependencies and behavior of the
  different components using \emph{Statika} and \emph{Datasets}
\item
  Parallelization and distributed analysis based on AWS, with on-demand
  infrastructure as the basic paradigm
\item
  Definition of complex pipelines using \emph{Loquat}, a composable
  system for scaling/parallelizing stateless computations especially
  designed for Amazon Web Services (AWS)
\item
  A new approach to data analysis specification, management and
  specification based on working with it in exactly the same way as for
  a software project, together with the extensive use of compile-time
  structures and checks
\item
  Modeling of the taxonomy tree using the new paradigm of graph
  databases (Bio4j). It facilitates the taxonomic assignment tasks and
  the calculation of the taxa abundance values considering the
  hierarchic structure of taxonomy tree (cumulative values)
\item
  per-read assignment (??)
\end{itemize}

>>>>

@@

Considering the current new metagenomics scenario and to tackle the
challenges posed by metagenomics big data analysis outlined in the
Introduction we have designed a new open source methodology for
analyzing metagenomics data. It exploits the new possibilities that
cloud computing offers to get a system robust, programmatically
configurable, modular, distributed, flexible, scalable and traceable in
which the biological databases of reference sequences can be easily
updated and/or frequently substituted by new ones or by databases
specifically designed for focused projects.

These are some of the more innovative MG7 features:

\begin{itemize}
\tightlist
\item
  Static reproducible specification of dependencies and behavior of the
  different components using \emph{Statika} and \emph{Datasets}
\item
  Parallelization and distributed analysis based on AWS, with on-demand
  infrastructure as the basic paradigm
\item
  Definition of complex workflows using \emph{Loquat}, a composable
  system for scaling/parallelizing stateless computations especially
  designed for Amazon Web Services (AWS)
\item
  A new approach to data analysis specification, management and
  specification based on working with it in exactly the same way as for
  a software project, together with the extensive use of compile-time
  structures and checks
\item
  Modeling of the taxonomy tree using the new paradigm of graph
  databases (Bio4j). It facilitates the taxonomic assignment tasks and
  the calculation of the taxa abundance values considering the
  hierarchic structure of taxonomy tree (cumulative values)
\item
  Exhaustive per-read taxonomic assignment using two complementary
  assignment algorithms Lowest Common Ancestor and Best BLAST Hit
\item
  Using a new 16S database of reference sequences (16S-DB7) with a
  flexible and sustainable system of updating and project-driven
  customization
\end{itemize}
